---
title: "Project1_Course5"
output: html_document
---
##1.Code for reading in the dataset and/or processing the data


```{r, echo= TRUE}
data <- read.csv("activity.csv", header = TRUE, sep = ",", na.strings = "NA")

```

###Summary of the data
```{r, echo= TRUE}
summary(data)
```  


###Str on data
```{r, echo= TRUE}
str(data)
```


###Head on data
```{r, echo= TRUE}
head(data)
```

###Convert date variable to date class
```{r, echo= TRUE}
data$date <- as.Date(data$date, format = "%Y-%m-%d")
str(data)
```

###Convert interval variable to factor
```{r, echo= TRUE}
data$interval <- factor(data$interval)
str(data)
```


##2- Histogram of the total number of steps taken each day

 
### Ignore missing value
```{r, echo= TRUE}
na_index <- is.na(as.character(data$steps))
data_no_NA <- data[!na_index,]
head(data_no_NA)
```


### Aggregating the number of steps taken each day:
```{r, echo= TRUE}
steps_each_day <- aggregate(steps ~ date, data = data_no_NA, sum)
colnames(steps_each_day) <- c("date", "steps")
head(steps_each_day)

```

###histogram of the total number of steps taken each day

```{r, echo= TRUE}
hist(as.numeric(steps_each_day$steps), breaks = 15, col = "pink", xlab = "Number of Steps", main= "Histogram of the total number of steps taken each day")

```

##3- Mean and median number of steps taken each day

###mean
```{r, echo= TRUE}
mean(steps_each_day$steps)

```

###Median
```{r, echo= TRUE}
median(steps_each_day$steps)

```


##4-Time series plot of the average number of steps taken


```{r, echo= TRUE}
steps_per_interval <- aggregate(data_no_NA$steps, by=list(interval=data_no_NA$interval), FUN=mean)
head(steps_per_interval)

#Adding column name
colnames(steps_per_interval) <- c("interval", "average_steps")

#ploting the average daily activity pattern 

plot(steps_per_interval$interval, steps_per_interval$average_steps, type = "l", xlab = "Interval", ylab = "Average Number of Steps", main = "Average Daily Activity Pattern", col = "blue")
```


##5-The 5-minute interval that, on average, contains the maximum number of steps

####The maximum number of average steps
```{r, echo= TRUE}
max_steps <- max(steps_per_interval$average_steps)
max_steps

```

####The 5-minute interval that contains the maximum number of steps
```{r, echo= TRUE}
intervale_max_steps<-steps_per_interval[which.max(steps_per_interval$average_steps),]$interval
intervale_max_steps

```

##6-Code to describe and show a strategy for imputing missing data

####The total number of missing values in the dataset (for each variable) - For "steps"
```{r, echo= TRUE}
sum(is.na(as.character(data$steps)))
```
#### For "date"
```{r, echo= TRUE}
sum(is.na(as.character(data$date)))
```
#### For "interval"
```{r, echo= TRUE}
sum(is.na(as.character(data$interval)))
```

#### Missing value is present in the "steps" variable. We can replace missing value by the mean of the 5 min interval.
```{r, echo= TRUE}
#finding the indices of missing values (NAs)
NA_index <- which(is.na(as.character(data$steps)))
complete_data <- data
#Imputing missing values using the mean for that 5-minute interval
complete_data[NA_index, ]$steps<-unlist(lapply(NA_index, FUN=function(NA_index){
                steps_per_interval[data[NA_index,]$interval==steps_per_interval$interval,]$average_steps
                }))

summary(complete_data)
str(complete_data)
```

##7-Histogram of the total number of steps taken each day after missing values are imputed

```{r, echo= TRUE}
steps_each_day_complete <- aggregate(steps ~ date, data = complete_data, sum)
colnames(steps_each_day_complete) <- c("date", "steps")

hist(as.numeric(steps_each_day_complete$steps), breaks = 20, col = "red", xlab = "Number of Steps", main= "Histogram of the total number of steps taken each day")

```


##8-Panel plot comparing the average number of steps taken per 5-minute interval across weekdays and weekends

```{r, echo= TRUE}
#Creating a factor variable "day "to store the day of the week:

complete_data$day <- as.factor(weekdays(complete_data$date))

#Creating a logical variable "is_weekday" (weekday=TRUE, weekend = FALE) :

complete_data$is_weekday <- ifelse(!(complete_data$day %in% c("Saturday","Sunday")), TRUE, FALSE) 

#Calculating the average number of steps for weekdays
weekdays_data <- complete_data[complete_data$is_weekday,]

steps_per_interval_weekdays <- aggregate(weekdays_data$steps, by=list(interval=weekdays_data$interval), FUN=mean)


#Calculating the average number of steps for weekends

weekends_data <- complete_data[!complete_data$is_weekday,]

steps_per_interval_weekends <- aggregate(weekends_data$steps, by=list(interval=weekends_data$interval), FUN=mean)


#Adding columns names
colnames(steps_per_interval_weekdays) <- c("interval", "average_steps")
colnames(steps_per_interval_weekends) <- c("interval", "average_steps")

#Adding a column to indecate the day
steps_per_interval_weekdays$day <- "Weekday"
steps_per_interval_weekends$day <- "Weekend"


#Merging the two togather
week_data <- rbind(steps_per_interval_weekends,steps_per_interval_weekdays)

#Converting the day variable to a factor
week_data$day <- as.factor(week_data$day)

#Making the plot
library(lattice)
xyplot(average_steps ~  interval | day, data = week_data, layout = c(1,2), type ="l", ylab="Number of Steps")


#The plot shows that that activity on the weekends tends to be more spread out over the day compared to the weekdays.


```

